### 2012-01-20 trunk/GUIErrorDetector/paper/technique.tex
revise implementation and evaluation sections

Given a sound call graph, our algorithm is sound in that it does not
miss true positives. We note, however, that computing a sound
call graph for partial programs such as an eclipse plugin,
or in the presence of reflection, is non-trivial. 
In particular, GUI applications built on some popular frameworks
often use configuration files and reflection to specify GUI layout
or extension points.
Existing call graph construction algorithms either suffer from incompleteness
or imprecision.  To alleviate this problem, 
We proposed a combined RTA
and k-CFA call graph construction algorithm in Section~\ref{sec:cg}.

We also note that a static analysis may report false positives due
to its conservative nature, or multiple warnings that acutally corrspond to
the same error. In addition, some GUI applications may use
Java native methods to interact with the underlying operating system.
Native methods are often beyond the ability of a static analysis, but
should be considered to construct a good call graph. To address the
above two limitations, we present a set of heuristics to filter
highly likely false positives and remove redundant reports, and provide
annotation support for programmers to specify call relationship for
 native methods.

@@

@@

@@

Unfortunately, GUI applications often use reflection, which makes either
RTA or k-CFA insufficient to build a satisfactory call graph. 
The example code in Figure~\ref{fig:sampleandroid}  illustrates their limitations.
In Figure~\ref{fig:sampleandroid}, line 7 uses reflection to create a \CodeIn{Button}
object by looking up its id declared in the XML file. Due to
its limitation in handling reflection, the k-CFA algorithm
will conclude that variable \CodeIn{button} at line 7 does not point to any
concrete \CodeIn{Button} object. Thus, it will miss the call edge \CodeIn{setText}
while processing line 11. On the other hand, a less precise algorithm like RTA~\cite{}
can not solve this problem. By explicitly specifying that GUI objects
like \CodeIn{Button} as declared in the XML file will be created, RTA maintains
a global value set for those GUI object type, and concludes that the \CodeIn{button} variable
at line 7 points to a concrete Button object. So the algorithm  adds a
call edge \CodeIn{setText} to the call graph. However, using RTA for
the whole program results in unnecessary precision loss. Note that,
RTA may conclude that the \CodeIn{cachedViews} variable at line 2
may point to any created \CodeIn{List} objects, despite the
fact that \CodeIn{cachedViews} can only be a \CodeIn{LinkedList} object at line 12.

To alleviate the above problems, we present a combined RTA and k-CFA
call graph construction algorithm as shown in Figure~\ref{fig:cgalgorithm}.
This algorithm uses the RTA algorithm to construct call graph for part of the program that
uses reflection to creating GUI objects (to ensure call graph completeness), while still uses
the k-CFA algorithm for the remaining part of the program (to avoid precision loss). Specifically,
users provide a list of classes, which can either be encoded manually or
automatically extracted from the XML configuration file, whose method invocations should be
constructed in a RTA way. The algorithm maintains a global set for
each  provided class, but a separate set for expressions of other
types.  For the example in Figure~\ref{fig:sampleandroid}, this
combined algorithm concludes that a \CodeIn{Button} object has been
created at line 7 and then adds the call edge \CodeIn{setText} to the
graph. It also concludes that the \CodeIn{cachedViews} variable can only
point to a \CodeIn{LinkedList} object, avoiding unnecessary precision loss.

%For desktop applications which have main method, it is straigtforward to apply
%the above algorithm. However, for applications on platforms like eclipse plugin
%and Android. presents new challenges. Two technical challenges must be solve.
%First, use reflection extensively, Second, it is possible to interact with
%native code.

@@

\subsubsection{Heuristic Filters and Annotations}
\label{sec:heuristic}

Static analysis exhaustively check possible error path; it may report
paths that indicate the same bug. The potentially huge volume of
reported warnings impose great burden for programmers to check
its validity.

a few useful static heuristic filtering rules.

\begin{itemize}
\item \textbf{Method call chain sumbsumption}. Remove redundant
results

\item \textbf{Remove system calls}. Method call chain involves typical
system methods like \CodeIn{toString} are unlikely to reveal a real bug.

\item \textbf{Method call chain with the same entry}. such
call chains may reveal the same bug

\item \textbf{Method call chain with the same tail}.  different ways
to reach the same bug

\item \textbf{User-defined rules}.  to overcome the limitation

The above filtering rules are not sound, it may filter real bugs but works
well in practice.

We also provide annotations for native methods which are beyond the ability
of most static analyses

@@

\item \textbf{Call graph entry points: } all overriden callback methods of the listerner classes.
Specifically, for each class, we check whether it is a Swing listerner class. If so,
we extract all its overriden callback methods as the call graph entry points.
 %Such methods are normally called by the eclipse framework
%by the UI thread. XXX This rule, though quite heuristical, works quite well in practice.

\item \textbf{UI accessing nodes:} unlike SWT, Swing does not explicitly check
whether the current thread is the UI thread or not, when a GUI object is accessed. Except
for three documented thread-safe method (\CodeIn{repaint}, XXXX), our technique
treat all methods defined in each GUI object as UI accessing nodes.

\item \textbf{Safe methods: } similar to SWT, Swing provides two utility methods
(\CodeIn{SwingUtilities.invokeLater} and \CodeIn{SwingUtilities.invokeDelay}) to pass messages to the UI thread. These two methods are regarded as safe methods.

>>>>

Given a sound call graph, our algorithm is sound in that it does not
miss true positives. We note, however, that computing a sound
call graph for partial programs such as an Eclipse plugin,
or in the presence of reflection, is non-trivial. 
In particular, GUI applications built on some popular frameworks
often use configuration files and reflection to specify GUI layout
or extension points.
Existing call graph construction algorithms suffer from incompleteness
, imprecision, or both.  To alleviate this problem, 
we proposed a combined RTA
and k-CFA call graph construction algorithm in Section~\ref{sec:cg}.

%we augument k-CFA with value set

We also note that a static analysis may report false positives due
to its conservative nature, or multiple warnings that actually correspond to
the same error. In addition, some GUI applications may use
Java native methods to interact with the underlying operating system.
Native methods are often beyond the ability of a static analysis, but
should be considered to construct a good call graph. To address the
above two limitations, we present a set of heuristics to filter
highly likely false positives and remove redundant reports, and provide
annotation support for programmers to specify call relationships for
 native methods.

@@

%In OO program, translate k-CFA as:
%k-call-site sensitive interprocedural pointer
%analysis with a k-context-sensitive heap and onthe-
%fly call-graph construction
%

@@

%this is wrong for k-CFA, k-CFA maintains several sets per expression, which k > 0
%should cite Ben Livshit's paper on "Reflection analysis for Java"
%Donot overclaim "say propose a new technique", instead, say "refine the technique"
%also need to cite Yannis's paper on solving k-CFA paradox

\textbf{need to describe why need to traverse from the UI entry, not just from the thread entry}

@@

Unfortunately, GUI applications often use reflection, which makes either
RTA or k-CFA insufficient to build a satisfactory call graph. 
The example code in Figure~\ref{fig:sampleandroid}  illustrates their limitations.
In Figure~\ref{fig:sampleandroid}, line 7 uses reflection to create a \CodeIn{Button}
object by looking up its id declared in the XML file. When this button
is clicked, its event handling code (line 10 -- 13) updates the text and caches
the current view to a list declared at line 2.

Existing call graph algorithms either suffer from the problem of incompleteness
or imprecision. The k-CFA algorithm performs points-to analysis with call
graph construction interleavingly, and maintains a value set for every expression.
For example, it concludes that variable \CodeIn{cachedViews} can only point
to a \CodeIn{LinkedList} object and the method call \CodeIn{add} at line 12
can only correspond to \CodeIn{LinkedList.add}. However, k-CFA fails to
conclude that variable \CodeIn{button} declared at line 7 points to
a \CodeIn{Button} object due to its limitation in dealing with reflection. Thus, k-CFA
may miss the call graph edge corresponding to the method call \CodeIn{setText} at line 11. 
On the other hand, using a coarse-grained algorithm like RTA~\cite{} can not still
fully solve this problem. RTA adds an edge for a method invocation 
to the call graph if an object of the receiver's type (or subtype)
has been created.  By explicitly specifying that all GUI classes declared in the XML file will be created,
RTA maintains a global value set for those GUI object type, and  adds a
call edge \CodeIn{setText} to the call graph. However, RTA introduces 
unnecessary precision loss: it may
may conclude that the \CodeIn{cachedViews} variable at line 2
point to any created \CodeIn{List} objects created somewhere else in the program, despite the
fact that \CodeIn{cachedViews} can only be a \CodeIn{LinkedList} object.

To alleviate the above problems, we present a combined RTA and k-CFA
call graph construction algorithm as shown in Figure~\ref{fig:cgalgorithm}.
The algorithm shares the same spirit as k-CFA to maintain a value set
for each expression. However, the only difference is that our
algorithm uses a RTA-style treatment for constructing the value set for
GUI classes: it parses the XML configuration file and explicitly
adds a raw object to each declared GUI variable's value set. Thus,
For the example in Figure~\ref{fig:sampleandroid}, this
combined algorithm concludes that a \CodeIn{Button} object has been
created at line 7 and then adds the call edge \CodeIn{setText} to the
graph. It also concludes that the \CodeIn{cachedViews} variable can only
point to a \CodeIn{LinkedList} object, avoiding unnecessary precision loss.
%For desktop applications which have main method, it is straigtforward to apply
%the above algorithm. However, for applications on platforms like eclipse plugin
%and Android. presents new challenges. Two technical challenges must be solve.
%First, use reflection extensively, Second, it is possible to interact with
%native code.

@@

\subsubsection{Heuristic Filtering and Annotation}
\label{sec:heuristic}

A static analysis can exhaustively check possible error path; but may report
paths that do not actually exist (i.e., false positives) or different paths
 that correspond to the same error (i.e., redundant warnings). We devise
a set of heuristics below to filter out likely false positives and redundant warnings.
The first two heuristics are for reducing false positives, while the
remaining ones are for removing redundant warnings.

%To remove those 
%The potentially huge volume of
%reported warnings impose great burden for programmers to check
%its validity.

\begin{enumerate}

\item \textbf{Filter Reports Containing Library Calls}. A
reported method call chain involving common Library calls
like \CodeIn{Object.toString} or \CodeIn{Runtime.shutDown} are unlikely to
be buggy, and should be filtered. In addition, a method call chain that \textit{only}
consists of library calls after the entry method is also unlikely
to reveal a valid bug, since it has no interaction with the UI. 
We encode a list of common library calls, and remove
a report if it contaiins one of them.
%Method call chain involves typical
%system methods like \CodeIn{toString} are unlikely to reveal a real bug.

\item \textbf{Filter Reports With User-Annotated Methods}. Besides general
library calls, users are also permitted
to explicitly annotate methods that will never trigger an error.
For example, the Android GUI framework provides a utility method 
\CodeIn{runOnUIThread} to check whether the current thread is the UI thread
before executing the code; if
not, it executes the code on the UI thread using message passing.
A reported call chain containing a annotated 
can be safely removed.
%method will never cause an
%invalid thread error and should be removed.

\item \textbf{Filter Lexically Redundant Reports}. One reported method call
chain can lexically subsume another one. For example, suppose that a reported method
call chain \CodeIn{a()} $\rightarrow$ \CodeIn{b()} $\rightarrow$ \CodeIn{c()}
leads to a potential error, and another entry method \CodeIn{d()} calls \CodeIn{a()}.
Therefore, a static analysis might also report another method call chain
\CodeIn{d()} $\rightarrow$ \CodeIn{a()} $\rightarrow$ \CodeIn{b()} $\rightarrow$ \CodeIn{c()}
that actually corresponds to the same error root as the previous one. In this case, 
 removing the second call chain would not reduce the error detection capability:
the first call chain already reveals the same error and is shorter 
for programmers to interpret.

\item \textbf{Filter Report With the Same $k$ Head Methods}. A method can call
multiple methods that access GUI objects, such as:
\begin{CodeOut}
\begin{alltt}
     public void m() \{
         accessUIObject1();
         accessUIObject2();
     \}
\end{alltt}
\end{CodeOut}
If method \CodeIn{m()} is invoked by a non-UI thread, two method call chains
only different at last few method nodes will be reported:

\CodeIn{a()} $\rightarrow$ ...\CodeIn{Thread.start()} ... $\rightarrow$ \CodeIn{m()} $\rightarrow$ \CodeIn{accessUIObject1()} ...

\CodeIn{a()} $\rightarrow$ ...\CodeIn{Thread.start()} ... $\rightarrow$ \CodeIn{m()} $\rightarrow$ \CodeIn{accessUIObject2()} ...

In fact, these two call chains correspond to the same error root: knowning
method \CodeIn{m()} is called by a non-UI thread is sufficient to understand and fix this error.
This filter compares every two reported method call chains that share the same
$k$ ($k$ is user-settable with a default value 5) head nodes, and removes the longer one.

\item \textbf{Filter Report With the Same $k$ Tail Methods}. A method
can have multiple callers, so that method call chains with the same tail are likely
to represent different ways to trigger the same error. This filter compares every
two reported method call chains that share the same 
$k$ ($k$ is user-settable with a default value 5) tail nodes, and removes the longer one.

\end{enumerate}

All above filters are based on heuristics, and thus not sound in
that it may filter real bugs. However, as we will show in
the next experiment section, all these filters work well, and significantly
reduces the number of false positives and redundant reports in practice.

A GUI application may also use native methods to interact with the underlying
operating system or platforms. However, native methods are often
beyond the ability of a static analysis but should be considered to make
the call graph more complete. To do so, we provide an annotation \CodeIn{@CallByNativeMethods}
for users to specify which native methods may call the current method. For example,
the following code snippet indicates that native methods \CodeIn{native1()} and \CodeIn{native2()}
will call method \CodeIn{javaMethod()}.

%\noindent 
\CodeIn{@CallByNativeMethods(callers=\{"native1", "native2"\})}

%\noindent
\CodeIn{public void javaMethod() \{ ... \}}

Our static analysis will take the call relationship as indicated by this
annotation into consideration when traversing the call graph. In our experiment,
we found such an annotation was useful: one real-world bug can only be reported
with the user-provided annotations.

@@

\item \textbf{Call graph entry nodes: } all overriden callback methods in the
listerner classes are called from the UI thread. We treat those callback
methods as the entry nodes.

%Specifically, for each class, we check whether
%it is a Swing listerner class. If so, we extract all its overriden callback methods as the call graph entry points.
 %Such methods are normally called by the eclipse framework
%by the UI thread. XXX This rule, though quite heuristical, works quite well in practice.

\item \textbf{UI accessing nodes:} as stated in Swing's documentation, methods
in Swing GUI classes are not thread safe with three exceptions (\CodeIn{repaint()},
\CodeIn{revalidate()}, and \CodeIn{invalidate()}). Thus, we
treat all methods as defined in each GUI class as UI accessing nodes.

\item \textbf{Safe methods: }  Swing provides two helper methods
(\CodeIn{SwingUti-\\lities.invokeLater} and \CodeIn{SwingUtilities.invokeAndWait}) to pass messages to the UI thread.
We treat these two methods as safe methods.
